---
title: "DBSCAN Outlier Detection"
author: "Simon Wills"
date: "2023-07-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Setup Environment
```{r}
library(devtools)
install_github("olobiolo/centroidr")
library(centroidr)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(centroidr,dbscan,fpc,jsonlite,shiny,GGally,WVPlots,ggplot2,SamSPECTRAL,ggrepel,ggpubr,gridExtra,mda)
```


## Now load up data to test for outliers.
## We need a data frame with a Specimen ID, Taxon (or grouping taxon) and the varibales to test for outliers
## The varibles could either be in 'raw' state or ordinated such as MDA, PCA or t-SNE
## I generally use MDA 

```{r}
df <- read.csv("C:/Users/swill/OneDrive - Natural History Museum/Manuscripts/PhD Chapters/MainThesis/CompleteThesis/DataForAppendicies/RESULTS_TrainingDatawithPostProbs_and_MDA.csv")

df <- as.data.frame(df[,c(2,3,46,47,48,49,50)])
```

## Clean up the data by dropping any clades with < 5 members
## There shouldn't be any as these should be dropped already.

```{r}
df <- df[df$GroupID %in% names(table(df$GroupID)) [table(df$GroupID) > 5],]
df$GroupID <- factor(df$GroupID)
```

```{r}
IDs <- sort(unique(df$GroupID))
```


## run DBSCAN

```{r}
plot_list_eps <- list()
counter = 0
for (V1 in IDs) {
  counter = counter +1
  cat(counter)
  kn.dist <- as.data.frame(sort(kNNdist(df[df$GroupID==V1,3:7],k=5)))
  kn.dist$point <- seq(1,nrow(kn.dist),by=1)
  colnames(kn.dist) <- c("dist","point")
  plot_list_eps[[V1]] <- ggplot(kn.dist) +
    geom_point(mapping=aes(x=point,y=dist),color="red") +
    theme(plot.title=element_text(size=8))+
    ggtitle(paste(V1))+
    labs(x="Point index") +
    labs(y="Distance (knn=5)") +
    theme_bw()
}

