---
title: "DeepTimePlotting"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: '2022-11-16'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


## setup environment
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(velociraptr,palaeoverse,deeptime,tidyverse,divDyn,ape,phytools,paleotree,dispRity,ggalt,dplyr,stringr,viridis,hrbrthemes,mapdata,cowplot,ggplot2,mapsf,sf,ggspatial,rnaturalearth,rnaturalearthdata,geojsonsf,RCurl,raster)
```

## to download and clean the data GetPBDBClean.Rmd script
## this produces a dataframe called taxa_filter



## get taxonomic ranges

```{r}
tax_range_time(occdf=taxa_filter,name="accepted_name", plot=TRUE)

taxa_range <- tax_range_time(occdf=taxa_filter,name="accepted_name", plot=FALSE)

```


# add useful columns
```{r}
# first add a 'genus' column based on the 'generic' part of taxon identifer
taxa_range$genus <- word(taxa_range$taxon,1)
#now extract the unique 'generic' names to a new df
unique_genus<- unique(c(as.character(taxa_range$genus)))
#create a numberic df based on the unique genus
taxa_range_colours<- data.frame(colours=as.numeric(factor(taxa_range$genus),levels=unique_genus))
# add it back to the main df
taxa_range$colours <- taxa_range_colours$colours
```


#generate time bins for plotting - assume full Mesozoic
```{r}

MesozoicPeriod_bins <- time_bins(interval = c("Mesozoic"),rank="period")
MesozoicEpoch_bins <- time_bins(interval = c("Mesozoic"),rank="epoch")
MesozoicStage_bins <- time_bins(interval = c("Mesozoic"),rank="stage")

MesozoicPeriod_Occs <- bin_time(taxa_filter,MesozoicPeriod_bins)
colnames(MesozoicPeriod_Occs)[colnames(MesozoicPeriod_Occs)=="bin_assignment"]="bin"
MesozoicPeriod_Occs <- merge(MesozoicPeriod_Occs,MesozoicPeriod_bins[,c("bin","interval_name","colour")],by="bin")


MesozoicEpoch_Occs <- bin_time(taxa_filter,MesozoicEpoch_bins)
colnames(MesozoicEpoch_Occs)[colnames(MesozoicEpoch_Occs)=="bin_assignment"]="bin"
MesozoicEpoch_Occs <- merge(MesozoicEpoch_Occs,MesozoicEpoch_bins[,c("bin","interval_name","colour")],by="bin")

MesozoicStage_Occs  <- bin_time(taxa_filter,MesozoicStage_bins) 
colnames(MesozoicStage_Occs)[colnames(MesozoicStage_Occs)=="bin_assignment"]="bin"
MesozoicStage_Occs <- merge(MesozoicStage_Occs,MesozoicStage_bins[,c("bin","interval_name","colour")],by="bin")

# summarise occurances by bin

MesozoicStageSummary <- MesozoicStage_Occs %>%
  group_by(bin_midpoint) %>%
  summarise(count=n())

# summary stage genera
MesozoicStageGeneraSummary <- MesozoicStage_Occs %>%
  group_by(bin_midpoint) %>%
  summarise(count=n_distinct(genus))


# or write to shapefile
MesEpoch_Shp <- MesozoicEpoch_Occs
coordinates(MesEpoch_Shp)=~lng+lat
shapefile(MesEpoch_Shp,"MesozoicEpochOccs.shp")

#create equal length bins for occ data - for diversity curve plots
# use bin_ages function
# https://psmits.github.io/paleo_book/managing-and-processing-data-from-the-paleobiology-database.html


# create 10 my bins in column bin_age

Mesozoic10ma_Occs <-
  taxa_filter %>%
  mutate(mid_ma = (max_ma + min_ma) / 2,
         bin = bin_ages(mid_ma, by = 10))

Mesozoic10ma_Occs <- 
  Mesozoic10ma_Occs %>%
  mutate(bin_age = bin_ages(mid_ma, by = 2, age = TRUE))

# summary 10 ma occurences
Mesozoic10maSummary <- Mesozoic10ma_Occs %>%
  group_by(bin_age) %>%
  summarise(count=n())

# summary 10 ma genera
Mesozoic10maGeneraSummary <- Mesozoic10ma_Occs %>%
  group_by(bin_age) %>%
  summarise(count=n_distinct(genus))
```

#Plot taxonomic range sorted by FAD
```{r}
# need to amend this plot so that the dumbell lines are coloured by taxon 

# reorder the taxa FAD oldest to youngest
taxa_range$taxon <- reorder(taxa_range$taxon,taxa_range$max_ma)
p <- ggplot(data=taxa_range,aes(x=max_ma,xend=min_ma,y=taxon,group=taxon))+
  geom_dumbbell(colour=taxa_range$colours, size=1) +
  scale_x_reverse() +
    theme_bw() +
  theme(axis.title.x = element_blank()) +
  labs(y="Taxon") +
  theme(axis.line = element_line(color='black'),
    plot.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank())


p <- gggeo_scale(p, dat = "periods", height = unit(1.5, "lines"), rot = 0, size = 2.5, abbrv = FALSE)

p <- gggeo_scale(p, dat = "epochs", height = unit(3, "lines"), rot = 90, size = 2.5, abbrv = TRUE)

p <- gggeo_scale(p, dat = "stages", height = unit(3, "lines"), rot = 90, size = 2.5, abbrv = TRUE)
```


#plot occurence data 10 ma bins
```{r}
p <- ggplot(Mesozoic10maSummary,aes(x=bin_age,y=count))+
 geom_line() +
 theme_bw() +
  scale_x_reverse() +
  theme(axis.title.x = element_blank()) +
  labs(y="Count")

p <- gggeo_scale(p, dat = "stages", height = unit(2, "lines"), rot = 90, size = 2.5, abbrv = TRUE)

pdf(file="10maBinOccs.pdf")
p
dev.off()

```

#oir just using the stage bins
```{r}
p <- ggplot(MesozoicStageSummary,aes(x=bin_midpoint,y=count))+
 geom_line() +
 theme_bw() +
  scale_x_reverse() +
  theme(axis.title.x = element_blank()) +
  labs(y="Count") 


p <- gggeo_scale(p, dat = "stages", height = unit(2, "lines"), rot = 90, size = 2.5, abbrv = TRUE)

pdf(file="StageBinOccs.pdf")
p
dev.off()

```

# Plot Genera Summary
```{r}
p <- ggplot(MesozoicStageGeneraSummary,aes(x=bin_midpoint,y=count))+
 geom_line() +
 theme_bw() +
  scale_x_reverse() +
  theme(axis.title.x = element_blank()) +
  labs(y="Count") 

p <- gggeo_scale(p, dat = "stages", height = unit(2, "lines"), rot = 90, size = 2.5, abbrv = TRUE)

pdf(file="StageBinGenera.pdf")
p
dev.off()

```


# plot both genera and occs on same plot
```{r}

SummaryGeneraOccs <-
  merge(MesozoicStageSummary,MesozoicStageGeneraSummary,by="bin_midpoint") %>% 
  rename(occs = count.x, genera = count.y)

scale = 10

p <- ggplot(SummaryGeneraOccs, aes(x = bin_midpoint, y=occs)) +
  geom_line(color = "black", size=2) +
  geom_line(aes(y=genera*10,color="OCC"),size=1) +
  labs(y = "Genera count per stage bin") +
  scale_y_continuous(sec.axis = sec_axis(~./10, name="OCC")) +
  coord_cartesian(xlim=c(195.05,70)) +
  scale_x_reverse() +
  theme_bw() +
  theme(axis.title.x = element_blank()) +
  theme(axis.title.y = element_blank()) +
  theme(legend.position="none") 


 p <-  gggeo_scale(p, dat = "stages", height = unit(2, "lines"), rot = 90, size = 2.5, abbrv = TRUE)
 
pdf(file="summarydiversity.pdf")
p
dev.off()
 
 
 

  
  




